Rotation and Lithium Confirmation of a 500 Parsec Halo for the Open Cluster NGC\,2516\footnote{This is paper 3 of the Cluster Difference Imaging Photometric Survey.}

### 2021-07-16 cluster_paper/ARXIV_VERSION/ms.tex
ngc2516: AJ accepted

\documentclass[12pt,twocolumn,tighten]{aastex63}
%\documentclass[12pt,twocolumn,tighten,linenumbers,trackchanges]{aastex63}
%\documentclass[12pt,twocolumn,tighten,linenumbers]{aastex63}
\usepackage{amsmath,amstext,amssymb}
\usepackage[T1]{fontenc}
\usepackage{apjfonts}
\usepackage[figure,figure*]{hypcap}
\usepackage{graphics,graphicx}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage[caption=false]{subfig} % for subfloat
\usepackage{enumitem} % for specific spacing of enumerate
\usepackage{epigraph}

@@

%\newcommand{\ncompstardenominator}{820} % plot_compstar_rotation.py
%\newcommand{\ncompstarnumerator}{312} % plot_compstar_rotation.py

@@

\newcommand{\kms}{\,km\,s$^{-1}$}
\newcommand{\ms}{\,m\,s$^{-1}$}
%\newcommand{\bpmrpo}{(\bpmrp)_0}
\newcommand{\bpmrpo}{(G_{\rm BP}-G_{\rm RP})_0}
\newcommand{\bpmrp}{G_{\rm BP}-G_{\rm RP}}

%% Reintroduced the \received and \accepted commands from AASTeX v5.2.
%% Add "Submitted to " argument.
\received{April 29, 2021}
\revised{July 7, 2021}
\accepted{---}
\submitjournal{AAS journals.}
\shorttitle{The Halo of NGC\,2516}

@@

% Cluster Difference Imaging Photometric Survey. III.
% Rotation and Lithium in an Open Cluster Spanning 500\,Parsecs
\title{
  Rotation and Lithium Confirmation of a 500 Parsec Halo for the Open
  Cluster NGC\,2516\footnote{This is paper 3 of the Cluster Difference
  Imaging Photometric Survey.}
}

@@

As part of this project, we focus here on a single southern open
cluster, and ask: is the cluster halo truly coeval with the core?  We
chose \cn\ for this analysis because it is young (100-200\,Myr) and
close enough ($d\approx400$\,pc) to facilitate \replaced{photometric
rotation measurements}{measurements of stellar rotation periods} using TESS.  \added{\citet{healy_stellar_2020}
in fact already used TESS rotation periods to study the stellar inclination
distribution of stars in the core of NGC\,2516.}
\replaced{Its halo}{The cluster's halo however} was reported by
\citet{kounkel_untangling_2019} to span
$\approx20\times10\times350$\,pc\added{, which is a larger volume than
has been considered in previous rotation analyses
\citep[][]{Irwin_NGC2516_2007,fritzewski_rotation_2020,healy_stellar_2020}}.
We want to know: is this halo real or is it just a collection of
interlopers, foreground and background stars?  To what extent can we
use Gaia alone to reliably identify age-dated needles in the haystack
of field stars?  What are the implications, observationally and
theoretically, if we can identify the dispersed halos of open
clusters?

\added{To assess whether the stars in the cluster's halo are the same
ages as stars in the cluster's core, we apply three different age-dating
techniques: isochrones, gyrochrones, and lithium depletion
\citep[see][for reviews]{soderblom_ages_2010,soderblom_ages_2014}.
Gyrochronology has now been empirically calibrated for FGK stars with
ages spanning $\approx$10\,Myr to $\approx$2.7\,Gyr \citep[{\it
e.g.},][]{rebull_usco_2018,curtis_temporary_2019,curtis_rup147_2020}.
Lithium equivalent widths can be used for relative age-dating of dwarf
stars over similar age ranges, though for a narrower range of spectral
types \citep{sestito_2005}.  One complication when deriving relative
lithium ages of the K dwarfs comes from the correlation between
lithium abundance and stellar rotation rate, which has recently been
reviewed by \citet{bouvier_lithium-rotation_2020}, and is a subject
that we explore in the latter portions of our analysis.  }

@@

    \vspace{-0.7cm}
		\subfloat{
			\includegraphics[width=0.700\textwidth]{f4b.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf TESS rotation periods and dereddened Gaia colors
  across the core and halo of \cn.} Sets $\mathcal{A}$ ({\it top})
  and $\mathcal{B}$ ({\it bottom}) undergo successive stages of
  automated cleaning \replaced{(see Section~\ref{subsubsec:cluster})}{; Set
  $\mathcal{B}$ is the same as Set $\mathcal{A}$, after selecting stars
  below the dashed line in the top panel (see Section~\ref{subsubsec:cluster})}.  The Pleiades
  \citep[125\,Myr;][]{rebull_rotation_2016a} and Praesepe
  \citep[650\,Myr;][]{douglas_poking_2017} are shown for reference.
  Stars in the core and the halo of NGC\,2516 overlap with the
  Pleiades at $\bpmrpo$$<$1.2.  From $1.2<\bpmrpo<1.7$, some NGC\,2516
  members have longer rotation periods than Pleiades members. Rotation
  periods for a comparison sample of field stars are shown in
  Appendix~\ref{app:compstar}.
  \label{fig:rot}
	}
\end{figure*}
\begin{figure}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.480\textwidth]{f5.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf Binarity indicators for rotators in \cn.} \replaced{Sets
			$\mathcal{A}$ ({\it top}) and $\mathcal{B}$ ({\it bottom}) are}{Set
			$\mathcal{B}$ is shown} as in
		Figure~\ref{fig:rot}.  Astrometric binaries ($\mathrm{RUWE}>1.2$;
		11\% of stars) appear in red; photometric binaries ($>$0.3\,mag
		above an empirical isochrone; 20\% of stars) are overplotted in
		orange.  
		Over $0.5<\bpmrpo<1.3$, Set $\mathcal{B}$ has \replaced{281}{523} stars.
		Dividing these stars into ``slow'' and ``fast'' sequences by eye,
		the fraction of stars showing signs of binarity in the slow and fast
		subsamples are \replaced{30\% (63/210)}{27\% (106/289)} and
		\replaced{56\% (40/71)}{51\% (68/134)}, respectively.
		\label{fig:binarity}
	}
\end{figure}

@@

We then measured the rotation period from the resulting light curve using
the \added{Lomb-Scargle} periodogram implementation\deleted{s} in
\texttt{astrobase}\replaced{, namely the traditional Lomb-Scargle, and
the Stellingwerf phase-dispersion minimization algorithm
\citep{lomb_1976,stellingwerf_period_1978,scargle_studies_1982,stellingwerf_period_2011,bhatti_astrobase_2018}}{ \citep{lomb_1976,scargle_studies_1982,bhatti_astrobase_2018}}.
We used the CDIPS aperture radius that, based on theoretical
expectations, was expected to give the optimal balance between light from
the target star and the sky background \citep{Sullivan_2015}.  This
typically resulted in a circular aperture radius of either 1 or 1.5
pixels.  We recorded the top five periodogram peaks\deleted{ from each
method}, and their corresponding powers.  Finally, as a check on
crowding, we recorded the number of stars within the aperture with
greater brightness than the target star, and with brightness within 1.25
and 2.5 TESS magnitudes of the target star.

The resulting rotation periods and periodogram powers are reported,
regardless of detection significance, in Table~\ref{tab:maintable}.
To clean these measurements, we designed 
automated cleaning criteria, which yielded sets we denote
$\mathcal{A}$ and $\mathcal{B}$.  For \replaced{both sets}{Set
$\mathcal{A}$}, we considered light
curves for which the peak Lomb-Scargle periodogram period was below 15
days, the normalized Lomb-Scargle power exceeded 0.08, and for which
no companions with brightness exceeding one-tenth of the target star
were in the aperture ({\it i.e.}, excluding visual binaries
with $\Delta T < 1.25$, according to the Gaia DR2 source catalog).
These limits were chosen after inspecting the light curves visually,
to eliminate low-quality rotation periods while retaining high
completeness.  For Set $\mathcal{A}$, this yielded \nautorotnumerator\
\replaced{light curves}{stars}, out of \nautorotdenominator\ stars \replaced{for which rotation
periods might plausibly have been detected}{that met the initial crowding requirements}.  For Set $\mathcal{B}$, we
additionally required that the Lomb-Scargle \replaced{and Stellingwerf
phase-dispersion methods yielded identical periods to within 10\%}{period
fell below the boundary drawn in Figure~\ref{fig:rot}, which was
constructed using the Pleiades polynomial from
\citet{curtis_rup147_2020}}.
This yielded \nautorotnumeratormatching\ light curves. 
Boolean flags for each set are included in Table~\ref{tab:maintable}.  A
set of field stars was also analyzed for comparison.  The results are
discussed in Appendix~\ref{app:compstar}.
\added{The rationale behind the additional cut in Set $\mathcal{B}$ is
that \cn\ is already known to have few if any stars above the slow
sequence \citep{fritzewski_rotation_2020}.  Stars above the slow sequence
are therefore most likely either not cluster members, or they are cluster
members, but with unresolved binary companions contaminating the rotation
period measurement \citep[{\it
e.g.},][Section~5.1]{stauffer_rotation_2016}.  For the purpose of
assessing which stars are rotationally consistent with being in the cluster,
this is a viable filter; for the purpose of exploring whether any
rotational outliers might be bonafide cluster members, Set $\mathcal{A}$
would be the better cut.}

@@

Figure~\ref{fig:rot} shows the resulting rotation periods, with Set
$\mathcal{A}$ shown on top and Set $\mathcal{B}$ on the bottom.  While
Set $\mathcal{A}$ is more complete than Set $\mathcal{B}$, this
completeness comes at the expense of greater contamination.  \deleted{For
instance, stars above the slow sequence (perhaps field contaminants)
mostly disappear in Set $\mathcal{B}$.  The ``void'' or ``gap''
underneath the slow sequence is also more pronounced in Set
$\mathcal{B}$.}  To facilitate a comparison against the Pleiades and
Praesepe, we used the rotation periods and dereddened colors from Table~5
of \citet{curtis_rup147_2020}, which drew on data from
\citet{rebull_rotation_2016a} and \citet{douglas_k2_2019} respectively.
The first order conclusion is that the Pleiades and \cn\ appear
gyrochronologically coeval for colors spanning $0.5<\bpmrpo<1.2$.  The
implications of this comparison for the age of \cn\ are discussed in
Section~\ref{disc:absage}.

\explain{Figure~\ref{fig:binarity} has been modified to account for the
revised definition of Set B.}

Figure~\ref{fig:binarity} shows the same data, with the points colored by
indicators of binarity.  To assess astrometric binarity, we used the
renormalized unit weight error (RUWE) from Gaia EDR3.  For plotting
purposes, we labeled anything with RUWE exceeding 1.2 as an astrometric
binary (11\% of the overall cluster sample; see {\it e.g.},
\citealt{belokurov_unresolved_2020}).  To assess photometric binarity, we
fitted a spline to Figure~\ref{fig:hr}, fixing the nodes by hand.  We
then labeled any points brighter than 0.3\,mag above the cluster sequence
as photometric binaries.  This included 20\% of the overall cluster
sample.  Many of the resulting candidate binaries, though not all of
them, appear below the slow rotation sequence.  In Set $\mathcal{B}$,
over the color range where the slow and fast sequence are present
($0.5<\bpmrpo<1.3$), there are \replaced{281}{523} stars, of which
\replaced{103 (37\%)}{174 (33\%)} show signs of binarity.  Dividing these
\replaced{281}{523} stars into ``fast'' and ``slow'' sequences by eye,
the fraction of stars showing signs of binarity in the slow and fast
subsamples are \replaced{30\% (63/210)}{27\% (106/289)} and \replaced{56\%
(40/71)}{51\% (68/134)}, respectively.  We speculate on possible
explanations in Section~\ref{disc:rotn}.

@@

	\end{center}
	\vspace{-0.7cm}
  \caption{ 
  {\bf \cn\ members exist hundreds of parsecs from the core, and up
  to $\approx$5\kms\ from the core in tangential velocity.} {\it Top
  left}: Cartesian positions (as in Figure~\ref{fig:XYZ}) and
  2D-tangential velocities for halo members.  Stars with detected
  rotation periods in \replaced{Set~$\mathcal{A}$}{Set~$\mathcal{B}$} are shown in blue;  halo
  members for which rotation periods should have been detectable, but
  were not detected, are shown in orange.  Stars in the core are shown
  as black points.  In
  the {\it top right}, the 
  ratio of detected to expected stars with rotation periods \added{in Set $\mathcal{B}$} is shown
  versus 3D separation from
  the core and 2D tangential velocity difference.  Bin widths are
  chosen to enforce the same number of stars in the denominator
  ($N=54$) per bin.  The {\it bottom panels} show the innermost and
  outermost 100 stars with detected rotation periods in position and
  velocity space\added{ (from Set~$\mathcal{A})$}.
  \label{fig:physical_x_rotn}
	}
\end{figure*}

@@

\explain{This paragraph contained a typo: the sentence containg ``at
least 25\,pc from the cluster core'' was meant to refer to the entire
cluster. }

The results are shown in the top two rows of
Figure~\ref{fig:physical_x_rotn}.  \added{Alternative visualizations
in the canoncial space of observed positions, parallaxes, and proper
motions are presented in Appendix~\ref{app:gaia6d_x_rotn}.}  In
\added{all} these figures, to ensure a fair comparison, we only show
stars with $0.5<\bpmrpo<1.2$ for which our TESS pipeline succeeded in
making light curves.  In other words, the stars in the base sample are
those for which we would have expected to detect a rotation period.
This color range is preferred because the the slow sequence becomes
less defined for spectral types later than $\approx$K4V.  The bins in
the histograms (top-right) were chosen to ensure that the same number
of stars were in each bin.  Rotation periods \added{consistent with a
gyrochronological age of 150\,Myr} are detected for
\replaced{66\%}{68\% (471/692)} of stars \replaced{at least 25\,pc
from the cluster core}{reported to be in the cluster}.  The average
detection rate inside of 25\,pc, \replaced{73\%}{72\% (290/404)}, is
slightly higher\added{ than outside of 25\,pc, where the detection
rate is 63\% (181/288)}. Within $\approx$3.5\,pc, the period detection
rate of \replaced{61\%}{59\% (26/44)} is
\replaced{relatively}{anomalously} low, due to crowding and saturation
in the TESS images.

The lower panels of Figure~\ref{fig:physical_x_rotn} give a second
view of how field star contamination affects the outskirts of the
cluster.  Of the outermost 100 stars in position space with detected
rotation periods\added{ in Set $\mathcal{A}$}, $\approx$8 appear as
outliers above the slow sequence, compared to just a few for the
innermost 100 stars.  For the outermost 100 stars in velocity space,
$\approx$13 appear as outliers.  After accounting for the stars that were
expected to show rotation periods and did not, this suggests field
contaminant rates of $\approx$50\% for the outermost cluster members in
our \added{initial} target list.

\replaced{Even with}{Despite} this level of contamination, the rotation period
measurements show that the halo extends to separations of
$\approx$250\,pc in physical space from the cluster core.  The most
widely separated F2V-K2V stars with rotation periods (blue points in
Figure~\ref{fig:physical_x_rotn}) are separated by $\approx$430\,pc.
The total length of the structure is therefore 400-500\,pc, depending
on which members of the halo are chosen as the ``tips'' on either end.
This agrees with the overall structure of the halo reported by
\citetalias{kounkel_untangling_2019}, with some minor caveats (orange
circles) visible in the top-left panels of
Figure~\ref{fig:physical_x_rotn}.

@@

The final approach we took for assessing the youth of the
halo population of \cn\ was an analysis of the Li\,\textsc{I}
6708\,\AA\ doublet.  \deleted{This type of age measurement has been reviewed by
\citet{soderblom_ages_2010}, and its connection to stellar rotation
has been discussed by \citet{bouvier_lithium-rotation_2020}.}
For \cn, two spectroscopic datasets seemed important:
Gaia-ESO \citep{gilmore_gaiaeso_2012} and GALAH
\citep{silva_galah_2015}.  At the time of analysis, Gaia-ESO DR4 and
GALAH DR3 were the most relevant
\citep{randich_gaiaeso_2018,buder_galah_2020}.  The target selection
and results from each survey were as follows.

@@

% the 21 stars that overlapped between GALAH and Gaia-ESO.  17
% of the 21 measurements agreed within 1-$\sigma$.  The remaining four
% were stars with $T_{\rm eff}>6200\,{\rm K}$, for which
% \citet{randich_gaiaeso_2018} reported Li EW detections ranging from
% $50$ to $75\,$m\AA, and for which we found either null results or only
% slightly significant ($\approx$30m\AA) detections.

% The outliers:
%    Fitted_Li_EW_mA     EWLi            source_id  Teff
%    24           -1.490   76.100  5290730488348194304  6786
%    34           19.485   71.300  5290935341107132800  6702
%    15           20.251   50.867  5290722310730727040  6197
%    25           29.378   52.500  5290724681552584192  6511

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.95\textwidth]{f8.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf Lithium and rotation in NGC\,2516.} Points show all
		candidate \cn\ members with TESS rotation periods\added{ (Set $\mathcal{A}$)} and Gaia-ESO or
		GALAH spectra, and are colored by equivalent width.  Points with
		reported equivalent widths below 20\,m\AA\ are shown with gray
		crosses.  The lithium-rotation correlation for the K-dwarfs has been
		observed in other young clusters, and is discussed in
		Section~\ref{discussion:lithium}.
		\label{fig:lithiumrot}
	}
\end{figure*}

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.95\textwidth]{f10.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf Comparison against previous analyses of stellar
  rotation in \cn.} 
  The bottom panel is a concatenation of Set $\mathcal{B}$ with the
  results of \replaced{\citet{Irwin_NGC2516_2007} (I07) and
  \citet{fritzewski_rotation_2020} (F20)}{\citet{Irwin_NGC2516_2007}
  (I07), \citet{fritzewski_rotation_2020} (F20), and
  \citet{healy_stellar_2020} (HM20)}.
  \replaced{I07 and F20}{I07, F20, and HM20} analyzed stars in the
  core of \cn.
  \label{fig:litcomp}
	}
\end{figure*}

\explain{Figure~\ref{fig:litcomp} has been updated to include rotation
periods from \citet{healy_stellar_2020}.}

>>>>

\documentclass[12pt,twocolumn,tighten]{aastex63}
\usepackage{amsmath,amstext,amssymb}
\usepackage[T1]{fontenc}
\usepackage{apjfonts}
\usepackage[figure,figure*]{hypcap}
\usepackage{graphics,graphicx}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage[caption=false]{subfig} % for subfloat
\usepackage{enumitem} % for specific spacing of enumerate
\usepackage{epigraph}

@@

@@

\newcommand{\kms}{\,km\,s$^{-1}$}
\newcommand{\ms}{\,m\,s$^{-1}$}
\newcommand{\bpmrpo}{(G_{\rm BP}-G_{\rm RP})_0}
\newcommand{\bpmrp}{G_{\rm BP}-G_{\rm RP}}

%% Reintroduced the \received and \accepted commands from AASTeX v5.2.
%% Add "Submitted to " argument.
\received{April 29, 2021}
\revised{July 7, 2021}
\accepted{July 16, 2021}
\submitjournal{AAS Journals}
\shorttitle{The Halo of NGC\,2516}

@@

\title{
  Rotation and Lithium Confirmation of a 500 Parsec Halo for the Open
  Cluster NGC\,2516\footnote{This is paper 3 of the Cluster Difference
  Imaging Photometric Survey.}
}

@@

As part of this project, we focus here on a single southern open
cluster, and ask: is the cluster halo truly coeval with the core?  We
chose \cn\ for this analysis because it is young (100-200\,Myr) and
close enough ($d\approx400$\,pc) to facilitate measurements of stellar rotation periods using TESS.  \citet{healy_stellar_2020}
in fact already used TESS rotation periods to study the stellar inclination
distribution of stars in the core of NGC\,2516.
The cluster's halo however was reported by
\citet{kounkel_untangling_2019} to span
$\approx20\times10\times350$\,pc, which is a larger volume than
has been considered in previous rotation analyses
\citep[][]{Irwin_NGC2516_2007,fritzewski_rotation_2020,healy_stellar_2020}.
We want to know: is this halo real or is it just a collection of
interlopers, foreground and background stars?  To what extent can we
use Gaia alone to reliably identify age-dated needles in the haystack
of field stars?  What are the implications, observationally and
theoretically, if we can identify the dispersed halos of open
clusters?

To assess whether the stars in the cluster's halo are the same
ages as stars in the cluster's core, we apply three different age-dating
techniques: isochrones, gyrochrones, and lithium depletion
\citep[see][for reviews]{soderblom_ages_2010,soderblom_ages_2014}.
Gyrochronology has now been empirically calibrated for FGK stars with
ages spanning $\approx$10\,Myr to $\approx$4\,Gyr \citep[{\it
e.g.},][]{rebull_usco_2018,curtis_temporary_2019,curtis_rup147_2020,barnes_rotation_2016}.
Lithium equivalent widths can be used for relative age-dating of dwarf
stars over similar age ranges, though for a narrower range of spectral
types \citep{sestito_2005}.  One complication when deriving relative
lithium ages of the K dwarfs comes from the correlation between
lithium abundance and stellar rotation rate, which has recently been
reviewed by \citet{bouvier_lithium-rotation_2020}, and is a subject
that we explore in the latter portions of our analysis.

@@

    \vspace{-0.7cm}
		\subfloat{
			\includegraphics[width=0.700\textwidth]{f4b.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf TESS rotation periods and dereddened Gaia colors
  across the core and halo of \cn.} Sets $\mathcal{A}$ ({\it top})
  and $\mathcal{B}$ ({\it bottom}) undergo successive stages of
  automated cleaning ; Set
  $\mathcal{B}$ is the same as Set $\mathcal{A}$, after selecting stars
  below the dashed line in the top panel (see Section~\ref{subsubsec:cluster}).  The Pleiades
  \citep[125\,Myr;][]{rebull_rotation_2016a} and Praesepe
  \citep[650\,Myr;][]{douglas_poking_2017} are shown for reference.
  Stars in the core and the halo of NGC\,2516 overlap with the
  Pleiades at $\bpmrpo$$<$1.2.  From $1.2<\bpmrpo<1.7$, some NGC\,2516
  members have longer rotation periods than Pleiades members. Rotation
  periods for a comparison sample of field stars are shown in
  Appendix~\ref{app:compstar}.
  \label{fig:rot}
	}
\end{figure*}
\begin{figure}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.480\textwidth]{f5.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf Binarity indicators for rotators in \cn.} Set
			$\mathcal{B}$ is shown as in
		Figure~\ref{fig:rot}.  Astrometric binaries ($\mathrm{RUWE}>1.2$;
		11\% of stars) appear in red; photometric binaries ($>$0.3\,mag
		above an empirical isochrone; 20\% of stars) are overplotted in
		orange.  
		Over $0.5<\bpmrpo<1.3$, Set $\mathcal{B}$ has 523 stars.
		Dividing these stars into ``slow'' and ``fast'' sequences by eye,
		the fraction of stars showing signs of binarity in the slow and fast
		subsamples are 27\% (106/289) and
		51\% (68/134), respectively.
		\label{fig:binarity}
	}
\end{figure}

@@

We then measured the rotation period from the resulting light curve using
the Lomb-Scargle periodogram implementation in
\texttt{astrobase} \citep{lomb_1976,scargle_studies_1982,bhatti_astrobase_2018}.
We used the CDIPS aperture radius that, based on theoretical
expectations, was expected to give the optimal balance between light from
the target star and the sky background \citep{Sullivan_2015}.  This
typically resulted in a circular aperture radius of either 1 or 1.5
pixels.  We recorded the top five periodogram peaks, and their corresponding powers.  Finally, as a check on
crowding, we recorded the number of stars within the aperture with
greater brightness than the target star, and with brightness within 1.25
and 2.5 TESS magnitudes of the target star.

The resulting rotation periods and periodogram powers are reported,
regardless of detection significance, in Table~\ref{tab:maintable}.
To clean these measurements, we designed 
automated cleaning criteria, which yielded sets we denote
$\mathcal{A}$ and $\mathcal{B}$.  For Set
$\mathcal{A}$, we considered light
curves for which the peak Lomb-Scargle periodogram period was below 15
days, the normalized Lomb-Scargle power exceeded 0.08, and for which
no companions with brightness exceeding one-tenth of the target star
were in the aperture ({\it i.e.}, excluding visual binaries
with $\Delta T < 1.25$, according to the Gaia DR2 source catalog).
These limits were chosen after inspecting the light curves visually,
to eliminate low-quality rotation periods while retaining high
completeness.  For Set $\mathcal{A}$, this yielded \nautorotnumerator\
stars, out of \nautorotdenominator\ stars that met the initial crowding requirements.  For Set $\mathcal{B}$, we
additionally required that the Lomb-Scargle period
fell below the boundary drawn in Figure~\ref{fig:rot}, which was
constructed using the Pleiades polynomial from
\citet{curtis_rup147_2020}.
This yielded \nautorotnumeratormatching\ light curves. 
Boolean flags for each set are included in Table~\ref{tab:maintable}.  A
set of field stars was also analyzed for comparison.  The results are
discussed in Appendix~\ref{app:compstar}.
The rationale behind the additional cut in Set $\mathcal{B}$ is
that \cn\ is already known to have few if any stars above the slow
sequence \citep{fritzewski_rotation_2020}.  Stars above the slow sequence
are therefore most likely either not cluster members, or they are cluster
members, but with unresolved binary companions contaminating the rotation
period measurement \citep[{\it
e.g.},][Section~5.1]{stauffer_rotation_2016}.  For the purpose of
assessing which stars are rotationally consistent with being in the cluster,
this is a viable filter; for the purpose of exploring whether any
rotational outliers might be bonafide cluster members, Set $\mathcal{A}$
would be the better cut.

@@

Figure~\ref{fig:rot} shows the resulting rotation periods, with Set
$\mathcal{A}$ shown on top and Set $\mathcal{B}$ on the bottom.  While
Set $\mathcal{A}$ is more complete than Set $\mathcal{B}$, this
completeness comes at the expense of greater contamination.    To facilitate a comparison against the Pleiades and
Praesepe, we used the rotation periods and dereddened colors from Table~5
of \citet{curtis_rup147_2020}, which drew on data from
\citet{rebull_rotation_2016a} and \citet{douglas_k2_2019} respectively.
The first order conclusion is that the Pleiades and \cn\ appear
gyrochronologically coeval for colors spanning $0.5<\bpmrpo<1.2$.  The
implications of this comparison for the age of \cn\ are discussed in
Section~\ref{disc:absage}.

Figure~\ref{fig:binarity} shows the same data, with the points colored by
indicators of binarity.  To assess astrometric binarity, we used the
renormalized unit weight error (RUWE) from Gaia EDR3.  For plotting
purposes, we labeled anything with RUWE exceeding 1.2 as an astrometric
binary (11\% of the overall cluster sample; see {\it e.g.},
\citealt{belokurov_unresolved_2020}).  To assess photometric binarity, we
fitted a spline to Figure~\ref{fig:hr}, fixing the nodes by hand.  We
then labeled any points brighter than 0.3\,mag above the cluster sequence
as photometric binaries.  This included 20\% of the overall cluster
sample.  Many of the resulting candidate binaries, though not all of
them, appear below the slow rotation sequence.  In Set $\mathcal{B}$,
over the color range where the slow and fast sequence are present
($0.5<\bpmrpo<1.3$), there are 523 stars, of which
174 (33\%) show signs of binarity.  Dividing these
523 stars into ``fast'' and ``slow'' sequences by eye,
the fraction of stars showing signs of binarity in the slow and fast
subsamples are 27\% (106/289) and 51\% (68/134), respectively.  We speculate on possible
explanations in Section~\ref{disc:rotn}.

@@

	\end{center}
	\vspace{-0.7cm}
  \caption{ 
  {\bf \cn\ members exist hundreds of parsecs from the core, and up
  to $\approx$5\kms\ from the core in tangential velocity.} {\it Top
  left}: Cartesian positions (as in Figure~\ref{fig:XYZ}) and
  2D-tangential velocities for halo members.  Stars with detected
  rotation periods in Set~$\mathcal{B}$ are shown in blue;  halo
  members for which rotation periods should have been detectable, but
  were not detected, are shown in orange.  Stars in the core are shown
  as black points.  In
  the {\it top right}, the 
  ratio of detected to expected stars with rotation periods in Set $\mathcal{B}$ is shown
  versus 3D separation from
  the core and 2D tangential velocity difference.  Bin widths are
  chosen to enforce the same number of stars in the denominator
  ($N=54$) per bin.  The {\it bottom panels} show the innermost and
  outermost 100 stars with detected rotation periods in position and
  velocity space (from Set~$\mathcal{A})$.
  \label{fig:physical_x_rotn}
	}
\end{figure*}

@@

The results are shown in the top two rows of
Figure~\ref{fig:physical_x_rotn}.  Alternative visualizations
in the canoncial space of observed positions, parallaxes, and proper
motions are presented in Appendix~\ref{app:gaia6d_x_rotn}.  In
all these figures, to ensure a fair comparison, we only show
stars with $0.5<\bpmrpo<1.2$ for which our TESS pipeline succeeded in
making light curves.  In other words, the stars in the base sample are
those for which we would have expected to detect a rotation period.
This color range is preferred because the the slow sequence becomes
less defined for spectral types later than $\approx$K4V.  The bins in
the histograms (top-right) were chosen to ensure that the same number
of stars were in each bin.  Rotation periods consistent with a
gyrochronological age of 150\,Myr are detected for
68\% (471/692) of stars reported to be in the cluster.  The average
detection rate inside of 25\,pc, 72\% (290/404), is
slightly higher than outside of 25\,pc, where the detection
rate is 63\% (181/288). Within $\approx$3.5\,pc, the period detection
rate of 59\% (26/44) is
anomalously low, due to crowding and saturation
in the TESS images.

The lower panels of Figure~\ref{fig:physical_x_rotn} give a second
view of how field star contamination affects the outskirts of the
cluster.  Of the outermost 100 stars in position space with detected
rotation periods in Set $\mathcal{A}$, $\approx$8 appear as
outliers above the slow sequence, compared to just a few for the
innermost 100 stars.  For the outermost 100 stars in velocity space,
$\approx$13 appear as outliers.  After accounting for the stars that were
expected to show rotation periods and did not, this suggests field
contaminant rates of $\approx$50\% for the outermost cluster members in
our initial target list.

Despite this level of contamination, the rotation period
measurements show that the halo extends to separations of
$\approx$250\,pc in physical space from the cluster core.  The most
widely separated F2V-K2V stars with rotation periods (blue points in
Figure~\ref{fig:physical_x_rotn}) are separated by $\approx$430\,pc.
The total length of the structure is therefore 400-500\,pc, depending
on which members of the halo are chosen as the ``tips'' on either end.
This agrees with the overall structure of the halo reported by
\citetalias{kounkel_untangling_2019}, with some minor caveats (orange
circles) visible in the top-left panels of
Figure~\ref{fig:physical_x_rotn}.

@@

The final approach we took for assessing the youth of the
halo population of \cn\ was an analysis of the Li\,\textsc{I}
6708\,\AA\ doublet.  
For \cn, two spectroscopic datasets seemed important:
Gaia-ESO \citep{gilmore_gaiaeso_2012} and GALAH
\citep{silva_galah_2015}.  At the time of analysis, Gaia-ESO DR4 and
GALAH DR3 were the most relevant
\citep{randich_gaiaeso_2018,buder_galah_2020}.  The target selection
and results from each survey were as follows.

@@

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.95\textwidth]{f8.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf Lithium and rotation in NGC\,2516.} Points show all
		candidate \cn\ members with TESS rotation periods (Set $\mathcal{A}$) and Gaia-ESO or
		GALAH spectra, and are colored by equivalent width.  Points with
		reported equivalent widths below 20\,m\AA\ are shown with gray
		crosses.  The lithium-rotation correlation for the K-dwarfs has been
		observed in other young clusters, and is discussed in
		Section~\ref{discussion:lithium}.
		\label{fig:lithiumrot}
	}
\end{figure*}

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.95\textwidth]{f10.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf Comparison against previous analyses of stellar
  rotation in \cn.} 
  The bottom panel is a concatenation of Set $\mathcal{B}$ with the
  results of \citet{Irwin_NGC2516_2007}
  (I07), \citet{fritzewski_rotation_2020} (F20), and
  \citet{healy_stellar_2020} (HM20).
  I07, F20, and HM20 analyzed stars in the
  core of \cn.
  \label{fig:litcomp}
	}
\end{figure*}

From their rotation period analysis of \cn,
\citet{fritzewski_rotation_2020} demonstrated that between 100 and
200\,Myr, young clusters (\cn, Blanco\,1, Psc-Eri, M\,35, M\,50, and
the Pleiades) all show similar rotation period distributions.  The FGK
stars show a ``slow sequence'', a population of ``rapid rotators'',
and in some cases stars in the ``void'' between the two populations
({\it e.g.,}, Psc-Eri and \cn\ show more stars in the void than the
Pleiades).  The late M dwarfs with detected rotation periods tend to
rotate rapidly ($<3$\,days), though there may also be late M dwarfs
with slow ($>20$\,days) rotation periods
\citep{stauffer_rotation_2016}.  Although these late M dwarfs are
beyond the magnitude limit of our analysis, they have been explored in
depth by \citet{Irwin_NGC2516_2007} and
\citet{fritzewski_rotation_2020}.  A comparison of our results is
shown in Figure~\ref{fig:litcomp}.  The bottom panel of this plot
includes, in order of precedence: rotation periods from our analysis,
from ``Class 1 and
2'' \citet{fritzewski_rotation_2020} sources, from
Table 1 of \citet{healy_stellar_2020}, and
from \citet{Irwin_NGC2516_2007}.  The main
contribution of our study beyond these previous analyses is our
sensitivity across both the core and halo of \cn, and a corresponding
expansion in the number of stars that can be analyzed.  In the
following section, we therefore focus on the origins of the features
in the rotation-color diagram, and whether our expanded sample can
provide any new insights.
